Untitled

### 2025-02-25 script/r/draft.Rmd
Add script precipitation, temperature (mean, max, min) + improved loops

print(levels, n=40)
```

#Adm setup

```{r}
cty_adm1 <- levels%>%
  filter(levels_nb>=1)%>%
  pull(country)

@@

tiff_file <- "data/data_raw/geography/desert/desert.tif"

@@

desert_r5_adm1 <- data.frame()

@@

  regions$desert_adm1 <- desert_adm1
  regions_df <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, desert_adm1) %>%
    mutate(round = 5)

  desert_r5_adm1 <- rbind(desert_r5_adm1, regions_df)
}

desert_r5_adm1 %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_r5_adm1))

saveRDS(desert_r5_adm1, file="data/data_rds/geo_ctrl/desert_r5_adm1.rds")

desert_r5_adm2 <- data.frame()

@@

  regions$desert_adm2 <- desert_adm2
  regions_df <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, desert_adm2) %>%
    mutate(round = 5)

  desert_r5_adm2 <- rbind(desert_r5_adm2, regions_df)
}

desert_r5_adm2 %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_r5_adm2))

saveRDS(desert_r5_adm2, file="data/data_rds/geo_ctrl/desert_r5_adm2.rds")
```

@@

elevation_r5_adm1 <- data.frame()

@@

  regions$elevation_adm1 <- elevation_adm1
  regions_df <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, elevation_adm1) %>%
    mutate(round = 5)

  elevation_r5_adm1 <- rbind(elevation_r5_adm1, regions_df)
}

elevation_r5_adm1 %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_r5_adm1))

saveRDS(elevation_r5_adm1, file="data/data_rds/geo_ctrl/elevation_r5_adm1.rds")

elevation_r5_adm2 <- data.frame()

@@

  regions$elevation_adm2 <- elevation_adm2
  regions_df <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, elevation_adm2) %>%
    mutate(round = 5)

  elevation_r5_adm2 <- rbind(elevation_r5_adm2, regions_df)
}

elevation_r5_adm2 %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_r5_adm2))

saveRDS(elevation_r5_adm2, file="data/data_rds/geo_ctrl/elevation_r5_adm2.rds")
```

>>>>

print(levels, n=40)

cty_adm1 <- levels%>%
  filter(levels_nb>=1)%>%
  pull(country)

@@

```{r}
tiff_file <- "data/data_raw/geography/desert/desert.tif"

@@

desert_adm1_df <- data.frame()

@@

  regions$desert_adm1 <- desert_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, desert_adm1)

  desert_adm1_df <- rbind(desert_adm1_df, country_results)
}

desert_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_adm1_df))

saveRDS(desert_adm1_df, file="data/data_rds/geo_ctrl/desert_adm1.rds")

desert_adm2_df <- data.frame()

@@

  regions$desert_adm2 <- desert_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, desert_adm2)

  desert_adm2_df <- rbind(desert_adm2_df, country_results)
}

desert_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(desert_adm2_df))

saveRDS(desert_adm2_df, file="data/data_rds/geo_ctrl/desert_adm2.rds")
```

@@

elevation_adm1_df <- data.frame()

@@

  regions$elevation_adm1 <- elevation_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, elevation_adm1)

  elevation_adm1_df <- rbind(elevation_adm1_df, country_results)
}

elevation_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_adm1_df))

saveRDS(elevation_adm1_df, file="data/data_rds/geo_ctrl/elevation_adm1.rds")

elevation_adm2_df <- data.frame()

@@

  regions$elevation_adm2 <- elevation_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, elevation_adm2)

  elevation_adm2_df <- rbind(elevation_adm2_df, country_results)
}

elevation_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(elevation_adm2_df))

saveRDS(elevation_adm2_df, file="data/data_rds/geo_ctrl/elevation_adm2.rds")
```

```{r}
library(viridis)

tiff_file <- "data/data_raw/geography/precipitation/CHELSA_bio12_1981-2010_V.2.1.tif"

precipitation_wgs84 <- rast(tiff_file)

print(precipitation_wgs84)

africa_extent <- ext(-20, 60, -40, 40)

precipitation_africa <- crop(precipitation_wgs84, africa_extent)

plot(precipitation_africa, main = "Precipitation Map of Africa", col = viridis(100), breaks = seq(0, 3000, by = 100))

precipitation_adm1_df <- data.frame()

  if (!identical(crs(precipitation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(precipitation_wgs84))
  }

  precipitation_wgs84_cropped <- crop(precipitation_wgs84, country_shp)
  precipitation_wgs84_masked <- mask(precipitation_wgs84_cropped, country_shp)

  precipitation_adm1 <- exactextractr::exact_extract(precipitation_wgs84_masked, regions, 'mean')

  regions$precipitation_adm1 <- precipitation_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, precipitation_adm1)

  precipitation_adm1_df <- rbind(precipitation_adm1_df, country_results)
}

precipitation_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(precipitation_adm1_df))

saveRDS(precipitation_adm1_df, file="data/data_rds/geo_ctrl/precipitation_adm1.rds")

precipitation_adm2_df <- data.frame()

  if (!identical(crs(precipitation_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(precipitation_wgs84))
  }

  precipitation_wgs84_cropped <- crop(precipitation_wgs84, country_shp)
  precipitation_wgs84_masked <- mask(precipitation_wgs84_cropped, country_shp)

  precipitation_adm2 <- exactextractr::exact_extract(precipitation_wgs84_masked, regions, 'mean')

  regions$precipitation_adm2 <- precipitation_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, precipitation_adm2)

  precipitation_adm2_df <- rbind(precipitation_adm2_df, country_results)
}

precipitation_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(precipitation_adm2_df))

saveRDS(precipitation_adm2_df, file="data/data_rds/geo_ctrl/precipitation_adm2.rds")
```

```{r}
shapefile_directory <- "data/data_raw/gadm/"

shapefile_names <- list.files(shapefile_directory, pattern = "*_2.shp", full.names = TRUE)

adm2_shapefiles <- lapply(shapefile_names, st_read)
adm2_combined <- do.call(rbind, adm2_shapefiles)

adm2_combined_merged <- merge(adm2_combined, precipitation_adm2, by.x = "NAME_2", by.y = "NAME_2", all.x = TRUE)

library(ggplot2)

ggplot() +
  geom_sf(data = adm2_combined_merged, aes(fill = precipitation_adm2), color = "black", size = 0.25) +
  scale_fill_viridis(name = "Precipitation (mm)", option = "plasma", na.value = "grey") +
  labs(title = "Precipitation Map of Africa",
       subtitle = "Regional Precipitation Values") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    legend.position = "bottom"
  )

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio1_1981-2010_V.2.1.tif"

temp_mean_wgs84 <- rast(tiff_file)

print(temp_mean_wgs84)

africa_extent <- ext(-20, 60, -40, 40)

temp_mean_africa <- crop(temp_mean_wgs84, africa_extent)

plot(temp_mean_africa, main = "temp_mean Map of Africa", col = rev(heat.colors(100)), breaks = seq(0, 50, by = 1))

temp_mean_adm1_df <- data.frame()

  if (!identical(crs(temp_mean_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_mean_wgs84))
  }

  temp_mean_wgs84_cropped <- crop(temp_mean_wgs84, country_shp)
  temp_mean_wgs84_masked <- mask(temp_mean_wgs84_cropped, country_shp)

  temp_mean_adm1 <- exactextractr::exact_extract(temp_mean_wgs84_masked, regions, 'mean')

  regions$temp_mean_adm1 <- temp_mean_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_mean_adm1)

  temp_mean_adm1_df <- rbind(temp_mean_adm1_df, country_results)
}

temp_mean_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_mean_adm1_df))

saveRDS(temp_mean_adm1_df, file="data/data_rds/geo_ctrl/temp_mean_adm1.rds")

temp_mean_adm2_df <- data.frame()

  if (!identical(crs(temp_mean_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_mean_wgs84))
  }

  temp_mean_wgs84_cropped <- crop(temp_mean_wgs84, country_shp)
  temp_mean_wgs84_masked <- mask(temp_mean_wgs84_cropped, country_shp)

  temp_mean_adm2 <- exactextractr::exact_extract(temp_mean_wgs84_masked, regions, 'mean')

  regions$temp_mean_adm2 <- temp_mean_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, temp_mean_adm2)

  temp_mean_adm2_df <- rbind(temp_mean_adm2_df, country_results)
}

temp_mean_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_mean_adm2_df))

saveRDS(temp_mean_adm2_df, file="data/data_rds/geo_ctrl/temp_mean_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio5_1981-2010_V.2.1.tif"

temp_max_wgs84 <- rast(tiff_file)

print(temp_max_wgs84)

africa_extent <- ext(-20, 60, -40, 40)

temp_max_africa <- crop(temp_max_wgs84, africa_extent)

plot(temp_max_africa, main = "temp_max Map of Africa", col = rev(heat.colors(100)), breaks = seq(0, 50, by = 1))

temp_max_adm1_df <- data.frame()

  if (!identical(crs(temp_max_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_max_wgs84))
  }

  temp_max_wgs84_cropped <- crop(temp_max_wgs84, country_shp)
  temp_max_wgs84_masked <- mask(temp_max_wgs84_cropped, country_shp)

  temp_max_adm1 <- exactextractr::exact_extract(temp_max_wgs84_masked, regions, 'mean')

  regions$temp_max_adm1 <- temp_max_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_max_adm1)

  temp_max_adm1_df <- rbind(temp_max_adm1_df, country_results)
}

temp_max_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_max_adm1_df))

saveRDS(temp_max_adm1_df, file="data/data_rds/geo_ctrl/temp_max_adm1.rds")

temp_max_adm2_df <- data.frame()

  if (!identical(crs(temp_max_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_max_wgs84))
  }

  temp_max_wgs84_cropped <- crop(temp_max_wgs84, country_shp)
  temp_max_wgs84_masked <- mask(temp_max_wgs84_cropped, country_shp)

  temp_max_adm2 <- exactextractr::exact_extract(temp_max_wgs84_masked, regions, 'mean')

  regions$temp_max_adm2 <- temp_max_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, temp_max_adm2)

  temp_max_adm2_df <- rbind(temp_max_adm2_df, country_results)
}

temp_max_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_max_adm2_df))

saveRDS(temp_max_adm2_df, file="data/data_rds/geo_ctrl/temp_max_adm2.rds")
```

```{r}
tiff_file <- "data/data_raw/geography/temperature/CHELSA_bio6_1981-2010_V.2.1.tif"

temp_min_wgs84 <- rast(tiff_file)

print(temp_min_wgs84)

africa_extent <- ext(-20, 60, -40, 40)

temp_min_africa <- crop(temp_min_wgs84, africa_extent)

plot(temp_min_africa, main = "temp_min Map of Africa", col = viridis(100), breaks = seq(-20, 30, by = 1))

temp_min_adm1_df <- data.frame()

  if (!identical(crs(temp_min_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_min_wgs84))
  }

  temp_min_wgs84_cropped <- crop(temp_min_wgs84, country_shp)
  temp_min_wgs84_masked <- mask(temp_min_wgs84_cropped, country_shp)

  temp_min_adm1 <- exactextractr::exact_extract(temp_min_wgs84_masked, regions, 'mean')

  regions$temp_min_adm1 <- temp_min_adm1

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_1, temp_min_adm1)

  temp_min_adm1_df <- rbind(temp_min_adm1_df, country_results)
}

temp_min_adm1_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_min_adm1_df))

saveRDS(temp_min_adm1_df, file="data/data_rds/geo_ctrl/temp_min_adm1.rds")

temp_min_adm2_df <- data.frame()

  if (!identical(crs(temp_min_wgs84), crs(country_shp))) {
    country_shp <- project(country_shp, crs(temp_min_wgs84))
  }

  temp_min_wgs84_cropped <- crop(temp_min_wgs84, country_shp)
  temp_min_wgs84_masked <- mask(temp_min_wgs84_cropped, country_shp)

  temp_min_adm2 <- exactextractr::exact_extract(temp_min_wgs84_masked, regions, 'mean')

  regions$temp_min_adm2 <- temp_min_adm2

  country_results <- regions %>%
    as.data.frame() %>%
    rename(iso = GID_0) %>%
    select(iso, NAME_2, temp_min_adm2)

  temp_min_adm2_df <- rbind(temp_min_adm2_df, country_results)
}

temp_min_adm2_df %<>%
  mutate(across(everything(), ~ ifelse(is.nan(.), NA, .)))

print(glimpse(temp_min_adm2_df))

saveRDS(temp_min_adm2_df, file="data/data_rds/geo_ctrl/temp_min_adm2.rds")
```
